\bf Title here

### 2020-08-20 paper/paper.Rmd
Section 4.2.1 searching plots

Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation. To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation.

@@

### Explore the value of objective function {#static}

@@

#### Searching points

A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimum progressively. We could plot the index value across its natural ordering, however, different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found. In the plot, points from those iterations towards the end will occupy the vast majority of the plot space. This motivates to use summarisation.  Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).

Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it does not report the number of points in each box. We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`. Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via a `cutoff` argument. A line geometry is also added to link the points with the largest index value in each iteration. This helps to visualise the improvement made in each iteration. Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as:

@@

##### Toy example: exploring searching points

We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour. Parameter `search_f = tour::search_better` and `max.tries = 500` is used. The index value of the searching points are shown in Figure \ref{points}. The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`). We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later.

```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs. The left plot does not use the plot space efficiently to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}

@@

dt %>% explore_trace_search()

@@

#### Interpolating points

@@

##### Toy example

@@

#### A toy example: understand different stage of search_geodesic

`search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space. Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions. The best one is picked and a line search is then run on the geodesic direction to find the target basis. The starting and target bases are then interpolated. In the next iteration, the target basis becomes the current basis and then the procedures continues.

```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}

>>>>

<!-- Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation. To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation. -->

@@

<!-- ### Explore the value of objective function {#static} -->

@@

### Explore searching points

A primary interest of diagnosing an optimisation algorithm is to study how it progressively finds its optimum. One way of doing it is to plot the index value across its natural ordering $t$, however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search. 
When plotting the searching observations as points in a plot, the space each iteartion take will be proportional to the number of points in that iteration.

<!-- different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found. In the plot, points from those iterations towards the end will occupy the vast majority of the plot space. This motivates to use summarisation.  Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).  -->

<!-- . We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`. Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via a `cutoff` argument. A line geometry is also added to link the points with the largest index value in each iteration. This helps to visualise the improvement made in each iteration. Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as: -->

Another option is to use summarisation for each iteration. Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis. This could be remedised by adding more layers using the concept of gramma of graphics. A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration. Further, a option to switch betwen displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations. This is avhieved via a `cutoff` parameter. Below we define the searching point plot based on four different component layers

@@

- Layer 4: label geom
  - data: A count table of the number of observation in each iteration
  - x: $j$ is mapped to the x-axis
  - y: the y-axis for each label is $0.99 * \text{MIN}$, where $\text{MIN}$ is the smallest index value in the whole data object
  - label: the number of observation is mapped to the label

<!-- *Example: exploring searching points* We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour. Parameter `search_f = tour::search_better` and `max.tries = 500` is used. The index value of the searching points are shown in Figure \ref{toy-search}. The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`). We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later. -->

Figure \ref{toy-search} presents a comparison between the two plotting options discussed above. Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasize the searches in the last few iterations. The second plot design evenly presents the summarised information for each iteration while allow for a switch to the full information for the iteration with small number of observations.

```{r toy-search, fig.height=5.5, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs. In upper plot, points from the last three iteration spans the vast majority of the plotting space leaving the first few iterationsbeing squeezed together whilst the lower plot spaces each iteration evenly and present summary information easy to read."}

@@

p1 <- dt %>% explore_trace_search()

p2 <- dt %>% ggplot() + 
  geom_point(aes(x = id, y = index_val, col = as.factor(tries))) + 
  theme(legend.position = "none")

p2 / p1

@@

### Explore interpolating points

@@

*Toy example* A toy example is shown in Figure \ref{toy-interp}.

```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
interp <- holes_2d_better_max_tries %>% 
  filter(info == "interpolation") %>% 
  mutate(id = row_number())

interp %>% 
  ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
  geom_point() + 
  geom_line() + 
  theme(legend.position = "bottom")
```

### Explore the searching space

@@

*A toy example: understand different stage of search_geodesic* `search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space. Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions. The best one is picked and a line search is then run on the geodesic direction to find the target basis. The starting and target bases are then interpolated. In the next iteration, the target basis becomes the current basis and then the procedures continues. A toy example is shown in Figure \ref{toy-pca}.

```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}

### 2020-08-20 paper/paper.tex
Section 4.2.1 searching plots

Since the global object is already tidy, not further tidying steps is
needed, while certain data wrangling steps
\citep{wickham2016rfordatascience} are still needed to transform the
global object into desirable format for one particular visualisation. To
emphasize on this good practice of data analysis, we will describe the
transformation steps needed for each diagnostic plots before stepping
into visualisation.

\hypertarget{static}{%
\subsubsection{Explore the value of objective function}\label{static}}

\hypertarget{searching-points}{%
\paragraph{Searching points}\label{searching-points}}

A primary interest of diagnosing an optimisation algorithm is to study
how it finds its optimum progressively. We could plot the index value
across its natural ordering, however, different iterations may have
different number of points and, towards the end of the search there
could easily be hundreds of bases being tested before the target basis
is found. In the plot, points from those iterations towards the end will
occupy the vast majority of the plot space. This motivates to use
summarisation. Rather than knowing the index value of \emph{every}
basis, we are more interested to have a general summary of all the index
value in that iteration and more importantly, the basis with the largest
index value (since it prescribes the next geodesic interpolation and
future searches).

Boxplot is a suitable candidate that provides five points summary of the
data, however it has one drawback: it does not report the number of
points in each box. We may risk losing information on how many points it
takes to find the target basis by displaying the boxplot alone for all
\texttt{tries}. Thus, the number of point in each iteration is displayed
at the bottom of each box and we provide options to switch iteration
with small number of points to a point geometry, which is achieved via a
\texttt{cutoff} argument. A line geometry is also added to link the
points with the largest index value in each iteration. This helps to
visualise the improvement made in each iteration. Using the concept of
\emph{gramma of graphics} \citep{wickham2010layered}, the plot for
exploring index value can be defined in three layers as:

@@

  \begin{itemize}
  \tightlist
  \item
    data: filter the points with the highest index value in group \(j\)
  \item
    x: \(j\) is mapped to the x-axis
  \item
    y: \(I\) is mapped to the y-axis
  \end{itemize}
\end{itemize}

\hypertarget{toy-example-exploring-searching-points}{%
\subparagraph{Toy example: exploring searching
points}\label{toy-example-exploring-searching-points}}

We choose variables \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
\texttt{x9} and \texttt{x10} to perform a 2D projection with tour.
Parameter \texttt{search\_f\ =\ tour::search\_better} and
\texttt{max.tries\ =\ 500} is used. The index value of the searching
points are shown in Figure \ref{points}. The label at the bottom
indicates the number of observations in each iteration and facilitates
the choice of \texttt{cutoff} argument (by default
\texttt{cutoff\ =\ 15}). We learn that the \texttt{search\_better}
quickly finds better projection basis with higher index value at first
and then takes longer to find a better one later.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/points-tries-1.pdf}
\caption{\label{points}A comparison of plotting the same search points
with different plot designs. The left plot does not use the plot space
efficiently to convey information from the plot while the right plot
provides good summarisation of data and number of points in each tries.}
\end{figure}

\hypertarget{interpolating-points}{%
\paragraph{Interpolating points}\label{interpolating-points}}

@@

\hypertarget{toy-example}{%
\subparagraph{Toy example}\label{toy-example}}

@@

\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
\paragraph{A toy example: understand different stage of
search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}

\texttt{search\_geodesic} is a two-stage ascending algorithm with four
different stages in the search and a PCA plot useful to understand how
the algorithm progresses and the relative position of each basis in the
PCA projected 2D space. Starting from the start basis, a directional
search is conducted in a narrow neighbourhood on five random directions.
The best one is picked and a line search is then run on the geodesic
direction to find the target basis. The starting and target bases are
then interpolated. In the next iteration, the target basis becomes the
current basis and then the procedures continues.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/pca-1.pdf}
\caption{\label{pca}PCA plot of search geodesic colouring by info allows
for better understanding of each stage in the geodesic search}
\end{figure}

@@

\hypertarget{explore-searching-space-1}{%
\subsubsection{Explore searching
space}\label{explore-searching-space-1}}

>>>>

\hypertarget{explore-searching-points}{%
\subsubsection{Explore searching
points}\label{explore-searching-points}}

A primary interest of diagnosing an optimisation algorithm is to study
how it progressively finds its optimum. One way of doing it is to plot
the index value across its natural ordering \(t\), however, it usually
takes the algorithm much longer to find a better basis than the current
one towards the end of the search. When plotting the searching
observations as points in a plot, the space each iteartion take will be
proportional to the number of points in that iteration.

Another option is to use summarisation for each iteration. Boxplot is a
suitable candidate that provides five points summary of the data,
however, there are two pieces of information missing from the boxplot:
1) It does not report the number of points, and 2) the position of the
last basis. This could be remedised by adding more layers using the
concept of gramma of graphics. A label geometry is added at the bottom
of the plot to show the number of points in each iteration and a line
geometry links the last basis in each iteration. Further, a option to
switch betwen displaying points and boxplot geometry is helpful since
point geometry can be more intuitive for the iteration with few
observations. This is avhieved via a \texttt{cutoff} parameter. Below we
define the searching point plot based on four different component layers

@@

  \begin{itemize}
  \tightlist
  \item
    data: filter the points with the highest index value in group \(j\)
  \item
    x: \(j\) is mapped to the x-axis
  \item
    y: \(I\) is mapped to the y-axis
  \end{itemize}
\item
  Layer 4: label geom

  \begin{itemize}
  \tightlist
  \item
    data: A count table of the number of observation in each iteration
  \item
    x: \(j\) is mapped to the x-axis
  \item
    y: the y-axis for each label is \(0.99 * \text{MIN}\), where
    \(\text{MIN}\) is the smallest index value in the whole data object
  \item
    label: the number of observation is mapped to the label
  \end{itemize}
\end{itemize}

Figure \ref{toy-search} presents a comparison between the two plotting
options discussed above. Using the natural order to plot searching
points distorts the point from the first few iterations and
over-emphasize the searches in the last few iterations. The second plot
design evenly presents the summarised information for each iteration
while allow for a switch to the full information for the iteration with
small number of observations.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-search-1.pdf}
\caption{\label{toy-search}A comparison of plotting the same search
points with different plot designs. In upper plot, points from the last
three iteration spans the vast majority of the plotting space leaving
the first few iterationsbeing squeezed together whilst the lower plot
spaces each iteration evenly and present summary information easy to
read.}
\end{figure}

\hypertarget{explore-interpolating-points}{%
\subsubsection{Explore interpolating
points}\label{explore-interpolating-points}}

@@

\emph{Toy example} A toy example is shown in Figure \ref{toy-interp}.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
\caption{\label{toy-interp} This is a toy example of plotting
interpolated points}
\end{figure}

\hypertarget{explore-the-searching-space}{%
\subsubsection{Explore the searching
space}\label{explore-the-searching-space}}

@@

\emph{A toy example: understand different stage of search\_geodesic}
\texttt{search\_geodesic} is a two-stage ascending algorithm with four
different stages in the search and a PCA plot useful to understand how
the algorithm progresses and the relative position of each basis in the
PCA projected 2D space. Starting from the start basis, a directional
search is conducted in a narrow neighbourhood on five random directions.
The best one is picked and a line search is then run on the geodesic
direction to find the target basis. The starting and target bases are
then interpolated. In the next iteration, the target basis becomes the
current basis and then the procedures continues. A toy example is shown
in Figure \ref{toy-pca}.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
\caption{\label{toy-pca}PCA plot of search geodesic colouring by info
allows for better understanding of each stage in the geodesic search}
\end{figure}

@@
